---
title: "SLA Analysis"
author: "Stephanie Pennington"
date: "5/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load necessary packages
library(pacman)

pacman::p_load(dplyr, tidyr, readr, knitr, kableExtra, lme4, car, sjPlot,
               lubridate, ggplot2, cowplot, ggmap, agricolae, broom, performance)

source("process_sla.R")

prettyp <- function(p.value) {
  if_else(p.value < 0.001, "P < 0.001", sprintf("P = %.3f", p.value))
}

# bootstrap options for kable
bo <- c("striped", "hover", "condensed", "responsive", full_width = FALSE)
```

# Motivation

Specific leaf area (SLA, the ratio of leaf area to leaf dry mass) is an important trait for plant physiological, structural, and modeling analyses. SLA is also a useful measurement to upscale plant production, vegetation community composition, and other processes in Earth System Models, providing current-day benchmarks and improving predictions of future vegetation structure and function in terrestrial ecosystems.

We investigate SLA differences between plots of varying elevation and salinity exposure along a tributary of the Chesapeake Bay. The goals of this project are to (i) quantify SLA of eight major tree species, at multiple canopy heights, in temperate deciduous forests at the Smithsonian Environmental Research Center (Maryland, USA); and (ii) test for effects of soil moisture and saltwater exposure on SLA by comparing samples taken at shoreline and upland positions, with the shoreline samples taken along a \~2 km tidal creek that provided a natural salinity gradient.

Map of SERC with plots labeled:

```{r serc-map, message=FALSE}
sites <- read.csv(file = "Design/sites.csv")
sites_bbox <- make_bbox(lon = sites$Longitude, lat = sites$Latitude, f = 0.8)
sites_map <- get_map(location = sites_bbox, source = "stamen", maptype = "terrain", zoom = 15)

ggmap(sites_map) +
  geom_point(data = sites, aes(x = Longitude, y = Latitude), color = "black", size = 1) +
  geom_text(data = sites, aes(x = Longitude, y = Latitude, label = Plot), size = 3, vjust = 2) +
  # Tower coordinates from https://www.neonscience.org/field-sites/serc
  annotate(geom = "point", x = -76.560014, y = 38.890131, size = 1) +
  annotate(geom = "text", x = -76.560014, y = 38.890131, label = "NEON Tower", size = 2.5, vjust = -1) +
  labs(x = "Longitude", y = "Latitude") +
  theme(axis.text = element_text(size = 10))

# ggsave("figures/fig1-map.png")
```

```{r tables, message=FALSE}

# Table 1: SLA by species
sla %>%
  group_by(Species_code, Species) %>%
  summarise(n = n(), mean_SLA = mean(SLA) %>% round(1),
    sd_SLA = sd(SLA), .groups = "drop") %>%
  kable(digits = 1, format = "html", caption = "Table 1: Specific leaf area (cm2/g) by species") %>% 
  kable_styling(bootstrap_options = bo)

# Table 2: SLA along salinity gradient
sla %>%
  filter(!is.na(Salinity)) %>%
  group_by(Species, Salinity) %>%
  summarise(
    mean_SLA = paste(round(mean(SLA), digits = 1), "±", round(sd(SLA), digits = 1),
      paste0("(", n(), ")")), .groups = "drop") %>%
  spread(Salinity, mean_SLA, fill = "") %>%
  kable(digits = 1, format = "html", caption = "Table 2: Specific leaf area (cm2/g) along salinity gradient") %>% 
  kable_styling(bootstrap_options = bo)

```

```{r upland-model, message = FALSE}

# Upland data only

# What controls SLA variability for upland (non-shoreline) trees?
# We're testing major fixed effects, Species and Position,
# as well as the interaction between Species and DBH, since there's a
# DBH-related SLA trend in at least some species.
# Finally, we account for overall plot-to-plot differences in SLA (for example
# due to a more fertile or sunny site) by including it as a random effect.
sla %>% 
  filter(Elevation == "Upland") -> sla_upland

lmer(SLA ~ Species + Species:DBH + Position + (1 | Plot),
     data = sla_upland) -> m_upland

# Pretty-print full model output
sjPlot::tab_model(m_upland)

# ANOVA table; this is what we report in the manuscript
car::Anova(m_upland, type = "III")
```

```{r shoreline-model, message = FALSE}

# Shoreline data only

sla %>% 
  filter(Elevation == "Shoreline") -> sla_shoreline

lm(SLA ~ Species * DBH + Position + Salinity,
     data = sla_shoreline) -> m_shoreline

```
